---
title: "Lecture 1: Social interactions, social outcomes, and social learning in social networks"
subtitle: "Getting started with computational social science modeling in R"
author: "Matthew A. Turner"
institution: "Department of Human Behavioral Sciences, Stanford Doerr School of Sustainability"
date: "1/6/2025"
format: revealjs
smaller: true
code-fold: true
fig-cap-location: bottom
    #html:
    #  fig-width: 8
    #  fig-height: 4
    #  code-fold: true
---


# Computational social science to promote sustainable behaviors


## Sustainable socio-ecological systems

![UN Sustainable Development Goals: [https://sdgs.un.org/goals](https://sdgs.un.org/goals)](1-resources/UN_SustDevGoals.png)

## Sustainable socio-ecological systems


![UN Sustainable Development Goals: [https://sdgs.un.org/goals](https://sdgs.un.org/goals)](1-resources/UN_SustDevGoals.png)


## Promoting sustainability needs computational social science



## Computational social _science_

- We use the scientific method, starting from the high school version:
    1. Hypothesize what causes some phenomenon 
    1. Gather evidence to support your explanation
    1. Evaluate the hypothesis (true, false, in-between)
    1. Repeat
- In computational social science we develop computer and mathematical models to analyze data and deduce the consequences of repeated social interactions subject to certain assumptions.

## The (computational social scientific) method

- We develop hypotheses 


# Diffusion of innovations

## How likely is the diffusion of an innovation?


:::: {.columns}

::: {.column width="65%"}
```{r}
library(ggplot2)
library(tibble)
#| label: fig-airquality
#| fig-cap: "Temperature and ozone level."
#| warning: false

ggplot(airquality, aes(Temp, Ozone)) + 
  geom_point() + 
  geom_smooth(method = "loess")
```
:::

::: {.column width="35%"}

- How many early-adopter _influencers_ must know a behavior in order to spread it?
- How does the perceived benefit of a behavior affect the spread of behavior?
:::

:::: 



# Social networks structure social interactions

## Social network construction

:::: {.columns}

```{r}
library(igraph)
igraph_options(vertex.color = "SkyBlue2", vertex.frame.color = "SkyBlue2", 
               vertex.label.cex = 4, vertex.size = 30, rescale = FALSE)
par(mar=c(0,0,0,0)+.1)
```
::: {.column width="65%"}
```{r}
#| fig-asp: 0.8
g <- make_graph(~ 1--2, 1--3, 2--3, 2--4, 4--5)
plot(g)
```
:::

::: {.column width=35%}
- A _vertice_ represents a person with a unique id
- An _edge_ represents a relationship between two people
- Two simulated people interact if and only if an edge connects them
- How many potential teachers does each person have?
:::

::::

## Social network properties
