---
title: "Coding vignettes to use in modeling social behavior."
format:
    html:
        toc: true
        echo: false
        fontsize: medium
bibliography: ../css4s.bib
---

These vignettes depend on some libraries:

```{r setup, include=FALSE}

# Load dependencies.
library(tibble)
library(dplyr)
library(ggplot2)
library(ggforce)
library(igraph)
library(forcats)
library(netrankr)
library(magrittr)

# Network illustration and igraph settings.
NETWORK_COLOR = "#69b3a2"
NODE_COLOR = adjustcolor(NETWORK_COLOR, alpha.f = 0.5)
igraph_options(vertex.frame.color = "white", vertex.color = NODE_COLOR, vertex.label.color = "black")

# Source the agent-based modeling code from socmod.R in root ProblemSets
# directory. Since this is in the CSS4S/notes directory for the course website
# I need to use a different path. Use `socmod_path` for the problem set.
socmod_path <- "~/workspace/CSS4S-ProblemSets/socmod.R" # <-- comment this out for your problem set 
# socmod_path <- "../socmod.R" ### <-- Use this one for your Problem Set
source(socmod_path)
```

# Plotting

## Networks


## Time series

See below for some examples of how to plot time series in R. Base R comes with plotting tools that work well for quick plots, and can be customized to make publication-ready plots. However, it is easier to make more complex plots with the tidyverse library `ggplot2`. 

For the examples to follow, use the following example output data from the example "random switching" model with some initial conditions to define ($N$ and $A_0 = A_{t=0}$, and one parameter, the probability an individual swtiches from one behavior to the other (`switch_prob`), which is identical for each behavior at each time step. 

```{r}
#tsteps <- 30
#series_default <- run_random_switch_example(tsteps)
#series_high_prob <- run_random_switch_example(tsteps, switch_prob)
```


Simple:
