---
title: "Computational Social Science for Sustainability"
format:
    html:
        toc: true
        echo: false
---

## Course notes {.unlisted}

# Foreword

These are course
notes for _Computational Social Science for Sustainability_.
This document is being regularly compiled, developed, and edited. 
Please send any feedback by [email](mailto:maturner@stanford.edu).

Last updated January 6, 2025.

# 1. Introduction (1/6/25 and 1/8/25)

If we are so sure that nudges, framing, and other psychological factors can
convince people to do one thing instead of another, why is it so difficult
sometimes to promote sustainable behavior? For many people, walking, biking, or using public transportation could save money and provide some exercise, yet public appetite for these modes of transportation is low and many cities in the USA seem permanently car-centric. In many cases electric vehicles would be an overall money-saver for individuals due to reduced energy and maintenance costs. Solar panels with batteries can power homes, and heat pumps can and regulate temperature in homes, more reliably than more expensive alternatives. So why don't more people choose these more sustainable and personally beneficial things?

It turns out that even a sustainable, highly-beneficial behavior can fail to
catch on simply because the social networks of who learns from whom have a
structure that inhibits the spread of _adaptive_ behaviors, i.e., those
behaviors that benefit an individual in a particular context. We can deduce
this fact using theory and models introduced in this course. But we can do more
to support decision making using the tools presented in this course. For
example, we can calculate the probability that an adaptation will be practiced
by every member of the model population, or if it will be abandoned by the
population and practiced by none. This serves as a proxy for measuring the
probability of _adaptive success_ for a population (all practice the adaptation
and not some alternative). Other computational tools introduced in this course
can predict how well different profit sharing strategies promote coperation for
sustainable use of common-pool resources. We can use these tools to explain why
healthy institutions can benefit both individuals and groups, and also how
polarization can paralyze institutions. These models begin from
empirically-motivated assumptions that fill in a general framework of social
and psychological mechanisms that generate emergent social phenomena including
the diffusion of adaptations, large-scale cooperation, and political
polarization.  <!--These models are implemented using a new R modeling
framework called [`socmod`]().-->

The initial use of social science for sustainability can be attributed to
Elinor Ostrom and her and colleagues' _socio-ecological systems_
framework [@Ostrom2009], for which she was awarded the Nobel prize in
economics. Socio-ecological systems was responding to the "zero contribution
thesis" that postulated "'unless there is coercion...rational, self-interested
individuals will not act to achieve their common or group interests'". But of
course humans do cooperate, to the point of regularly risking their own lives
to save the lives of others, and more quotidien cooperation for roads, public
safety, national security, just to name a few. (HOW DID THEY RESPOND? WHAT'S THE ANSWER?)

Ostrom's eight design principles for sustainable natural resource management have especially elevated Indigenous adaptations that had been ignored since colonization. 
(EXAMPLES)
This is especially notable for our purposes because it forces us to abandon the word "innovation" when it comes to sustainable behaviors. 

We want to build on Ostrom's progress by developing a scientific modeling framework
to represent key mechanisms that constrain and drive social transmission. This framework is designed to generalize three types of social interaction: (1) common-pool resource dilemmas and solutions treated by Ostrom; (2) non-cooperative personal behavior decisions like mode of commute or home efficiency improvements; and (3) social influence of opinions that can result in consensus, extremism, or polarization depending on context including and especially the social network that structures interactions. 

## Social networks structure social interaction

Social interactions here are always structured by social networks, starting with
our tutorial example, "The Influencer Dynamic", below. Social networks are a mechanical representation of a relationship between two people. In networks, people are represented as _nodes_ or _vertices_, which are typically represented by a circle or another
shape. Relationships between people are represented as lines connecting nodes
called _edges_. In our social modeling framework, networks are real-world instances 
of graphs, the mathematical collection of nodes with edges represented numerically by an
_adjacency matrix_. (Continue introducing social networks...)

```{r}
#| output: false
#| echo: false
library(ggforce)
library(tibble)
```


```{r}
#| fig-width: 5.5
#| fig-asp: 0.33333
#| label: fig-teacher-learner-infoflow
#| fig-cap: "Social networks represent relationships between people, with arrows pointing in the direction of information flow, representing who learns from whom."
circles <- tibble(x0 = c(0, 2), y0 = c(0, 0), r = c(.5, .5))
ggplot() + 
    geom_circle(aes(x0 = x0, y0 = y0, r = r), data = circles, fill = "SkyBlue2") + 
    geom_segment(aes(x = 0.5, y = 0, xend = 1.5, yend = 0), arrow = arrow()) +
    annotate("text", x=0, y=0, label="Teacher", size=8) +
    annotate("text", x=2, y=0, label="Learner", size=8) +
    annotate("text", x=0.95, y=0.15, label="Information flow", size=5) +
    theme_void()
```


(TODO: add bi-directional arrows and mention equivalent to a line with no arrows)



# Diffusion of adaptations via social learning

To begin, consider the simple case of people learning from an influencer, ignoring social learning from others.

## The Influencer Dynamic (1/8 and reviewed 1/13)



## Innovative adaptations
